---
title: "validR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{validR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The validR is a set of functions to be used for testing and validating data. 
The main functions are:

- test_range
- test_values
- test_unique
- test_na
- test_bool (under development)

To use the testing functions first create the test setup. How each test class should be setup can be found by ?class_test_range in the details section.

- class_test_range
- class_test_values
- class_test_unique
- class_test_na
- class_test_bool  (under development)


```{r, eval = FALSE, echo = TRUE}
library(validR)

df <- data.frame(x = 1:10, y = c(rep("X", 4), rep("Y", 4), rep("Z", 2)))


setup1 <- setup_test_range(df_name = "df", col_name = "x", int = TRUE, lower_inclu = TRUE,
                           upper_inclu = TRUE, lower = 1, upper = 10, na = FALSE)

setup2 <- setup_test_values(df_name = "df", col_name = "y", values = c("X", "Y"), na = FALSE)

setup3 <- setup_test_unique(df_name = "df", col_name = "x", na = FALSE)

setup4 <- setup_test_na(df_name = "df", col_name = "y")

test1 <- apply_test(setup1, df)
test2 <- apply_test(setup2, df)
test3 <- apply_test(setup3, df)
test4 <- apply_test(setup4, df)

tests <- rbind(test1, test2, test3, test4)

```